---
title: "Final Project Data Memo"
author: "Jamie Lee, Susan Tran, Grace Park"
date: "5/7/2021"
output: 
  html_document:
    toc: true
    toc_float: true
    highlight: "tango"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Source

For our final project, our group has decided to use data from Mac Miller's discography. We have retrieved two datasets from \_\_\_\_\_\_\_. One dataset consists of lyrical data from Mac Miller's songs and the other dataset contains word-by-word data.

```{r}
malcolm <- readRDS("data/processed/malcolm.rds") # lyrical data
macaroni <- readRDS("data/processed/macaroni.rds") # word-by-word data

skimr::skim_without_charts(malcolm)
skimr::skim_without_charts(macaroni)
```

From our initial skimming of the datasets, we can see that there are no major missingness issues, although 8 values under the "lyric" variable are missing in the "malcolm" dataset and 8 values under the "word" variable are missing in the "macaroni" dataset.

<br>

```{r}
malcolm %>% filter(is.na(lyric))
macaroni %>% filter(is.na(word))
```

<br>

Taking a closer look at the missingness shows that, as expected, it is of the same lines across both datasets, and it seems that a lot of this missingness is from the first line of the songs, though not all.

<br>

## Why this Data

As we are all fans of Mac Miller's work, our group wanted to explore

## Visualizing the Data

## Potential Data Issues
